### 2021-04-09 ch05_paper_ecs/paper_ecs.tex
Used lowercase versions of namecref(s) at correct places

The massive spread in \ac{ECS} within the \acs{CMIP}6 ensemble discussed in the
previous \namecref{ch:04:papers_ecs_tcr_assessment} begs the question about the
interpretation and credibility of these results. Since the \acs{CMIP}6 models
provide crucial input for the upcoming international climate assessment of the
\ac{IPCC} (\acs{AR}6), this is not only important for the scientific community,
but also directly affects policymakers and society. Due to the many new and
improved processes included to the models, the \acs{CMIP}6 ensemble is arguably
more realistic than its predecessor generations. However, does this
automatically make results derived from \acs{CMIP}6 models more trustworthy
than those derived from earlier \ac{CMIP} models? In general, the answer to
this question is \enquote{not necessarily}. Even the \acs{CMIP}6 models are far
from fully simulating all aspects and processes of the full vastly complex
Earth system. Therefore, the addition of new processes to these models could
introduce new biases which have not been present in earlier \ac{CMIP} models.
For example, the apparent increase in climate sensitivity in the \acs{CMIP}6
models might be wrongly caused by missing processes that cancel out the effects
of the newly included processes. Thus, it is vital to take other sources of
information into account to assess the credibility of \ac{ESM} results.

@@

\begin{table}[!t]
  \centering
  \hbadness=3500
  \begin{tabular}{p{0.07\columnwidth} p{0.24\columnwidth} p{0.29\columnwidth}
      p{0.29\columnwidth}}
    \toprule
    Label & Reference & Description of \xaxis{} & Variables and corresponding
    observations \\
    \midrule
    BRI & \textcite{Brient2016} & Sensitivity of shortwave cloud albedo
    to changes in \acf{SST} [$\si{\pctK{}}$] & \makecell{\tabitem\emph{hur}
      (ERA-Interim) \\ \tabitem \emph{rsdt} (CERES-EBAF) \\ \tabitem
      \emph{rsut} (CERES-EBAF) \\ \tabitem \emph{rsutcs} (CERES-EBAF) \\
      \tabitem \emph{ts} (HadISST)} \\
    \midrule
    COX & \textcite{Cox2018} & $\Psi$ (temperature variability metric)
    [$\si{K}$] &
    \tabitem \emph{tas} (HadCRUT4) \\
    \midrule
    LIP & \textcite{Lipat2017} & Extent of Southern Hemisphere Hadley cell
    [$\si{\degree}$] & \tabitem \emph{va} (ERA-Interim) \\
    \midrule
    SHD & \textcite{Sherwood2014} & $D$-index (large-scale lower-tropospheric
    mixing) [$\si{1}$] & \tabitem \emph{wap} (ERA-Interim) \\
    \midrule
    SHL & \textcite{Sherwood2014} & \acs{LTMI} (\acl{LTMI}) [$\si{1}$] &
    \makecell{\tabitem \emph{hur} (ERA-Interim) \\ \tabitem \emph{ta}
      (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \midrule
    SHS & \textcite{Sherwood2014} & $S$-index (small-scale lower-tropospheric
    mixing) [$\si{1}$] & \makecell{\tabitem \emph{hur} (ERA-Interim) \\
      \tabitem \emph{ta} (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \midrule
    SU & \textcite{Su2014} & Error in vertical profile of relative humidity
    [$\si{1}$] & \tabitem \emph{hur} (AIRS, MLS-Aura) \\
    \midrule
    TIH & \textcite{Tian2015} & Tropical mid-tropospheric humidity asymmetry
    index from AIRS [$\si{\%}$] & \tabitem \emph{hus} (AIRS) \\
    \midrule
    TII & \textcite{Tian2015} & Southern \acs{ITCZ} index from GPCP
    [$\si{\mmday{}}$] & \tabitem \emph{pr} (GPCP) \\
    \midrule
    VOL* & \textcite{Volodin2008} & Difference between tropical and southern
    midlatitudinal cloud fraction [$\si{\%}$] & \tabitem \emph{clt} (ISCCP
    D-2) \\
    \midrule
    ZHA* & \textcite{Zhai2015} & Seasonal response of \acl{MBLC} fraction to
    changes in \acs{SST} [$\si{\pctK{}}$] &
    \makecell{\tabitem \emph{cl} (CloudSat/CALIPSO) \\ \tabitem \emph{tos}
      (AMSRE \acs{SST}) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \bottomrule
  \end{tabular}
  \caption[
    Overview of the 11 emergent constraints on the \acf{ECS} evaluated in
    \cref{ch:05:paper_ecs}.
  ]{
    Overview of the 11 emergent constraints on the \acf{ECS} evaluated in this
    \namecref{ch:05:paper_ecs}. Detailed descriptions of the variables used to
    calculate the \xaxis{} values of the emergent constraints and the
    references for the corresponding observations (fourth column) are given in
    \cref{tab:app:a:overview_variables,tab:app:a:observations}, respectively.
    For emergent constraints marked with an asterisk (*), the observational
    value of the original publication is used to calculate the observational
    constraint. For all others, the observational value is computed with the
    \acs{ESMValTool}. \AdaptedFrom{Schlund2020a}.
  }
  \label{tab:05:overview_emergent_constraints}
\end{table}

This work is already published in \textcite{Schlund2020a}. For this paper, the
author of this thesis led the writing and the analysis and implemented the code
to reproduce this study with all figures and tables using the \ac{ESMValTool}.
In this \namecref{ch:05:paper_ecs}, we first introduce the data (climate model
output and observations) and methods used in this study
(\cref{sec:05:data_and_methods}). After that, we present the 11 emergent
constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5 and \acs{CMIP}6
ensemble (\cref{sec:05:comparison_of_emergent_constraints}). Finally, we
provide a discussion (\cref{sec:05:discussion}) and a summary of the results
(\cref{sec:05:summary}).

@@

In this \namecref{sec:05:comparison_of_emergent_constraints} we describe and
discuss the 11 emergent constraints on \ac{ECS} summarized in
\cref{tab:05:overview_emergent_constraints} using \acs{CMIP}5 and \acs{CMIP}6
data
(\cref{subsec:05:bri,subsec:05:cox,subsec:05:lip,subsec:05:shd,subsec:05:shs,%
  subsec:05:shl,subsec:05:su,subsec:05:tih,subsec:05:tii,subsec:05:vol,%
  subsec:05:zha}) and provide a best estimate for \ac{ECS} and statistical
significance of the 11 emergent constraints in
\cref{subsec:05:emergent_constraints_summary}. While most of these emergent
constraints have been derived using data from the \acs{CMIP}5 and/or
\acs{CMIP}3 ensembles, to our knowledge none of them has been evaluated on
the \acs{CMIP}6 ensemble so far. The results for the individual emergent
constraints described in the following are shown in
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}.
The left columns in these figures show the emergent relationships, including
the uncertainty of the linear regressions (blue and orange shaded areas; see
\cref{eq:02:spe}) and the uncertainty in the observations (gray shaded area;
see \cref{eq:02:pdf_obs}). The right columns show the \acp{PDF} of \ac{ECS} in
the original model ensemble (histogram) and the constrained distribution given
by the emergent constraints (blue and orange line; see
\cref{eq:02:pdf_y_given_x0}). \Cref{tab:05:overview_results} shows the
corresponding $66 \unit{\%}$ confidence intervals (\ie{} the
$\rangeunit{17}{83}{\%}$ intervals) of \ac{ECS} derived from the \acp{PDF}
given by \cref{eq:02:pdf_y_given_x0} and the $p$-values used to assess the
significance of the emergent relationships.

@@

As shown in the previous \namecrefs{sec:05:comparison_of_emergent_constraints},
most emergent relationships show smaller coefficients of determination when
evaluated on the new \acs{CMIP}6 ensemble compared to the \acs{CMIP}5 ensemble.
In this \namecref{sec:05:discussion}, we discuss possible reasons for, and
implications of, these differences. As reported by \textcite{Caldwell2014}, the
large amount of data provided by modern \acp{ESM} can generate spurious
correlations of variables between past climate and \ac{ECS} just by chance,
especially when only a few climate models are considered. This would cause the
performance of the emergent constraint to be reduced on out-of-sample data
(like the new \acs{CMIP}6 ensemble), since the emergent relationship appeared
just by chance and not because of a physically based mechanism.

@@

In this \namecref{ch:05:paper_ecs}, 11 different emergent constraints on
\ac{ECS}, which are mostly directly or indirectly related to cloud feedbacks,
are assessed by applying them to results from \acp{ESM} contributing to
\acs{CMIP}5 and \acs{CMIP}6. Of particular interest are the results from
\acs{CMIP}6, since all analyzed emergent constraints have been published prior
to the availability of \acs{CMIP}6 data. In summary, the best estimate of
\ac{ECS} averaged over all emergent constraints increases by $12 \unit{\%}$
when moving from relationships trained on \acs{CMIP}5 to those trained on
\acs{CMIP}6. Some increase is predicted by almost every constraint we analyze
(with TII as the only exception) and can at least be partly explained by the
increased multi-model mean \ac{ECS} of \acs{CMIP}6, which has not been
accompanied by systematic changes in the constraint variables that could
explain this increase, leading to regression fits with higher intercept values
at observed constraint values. This is also illustrated by the \acs{CMIP}5 and
\acs{CMIP}6 \acp{MMM} in the left columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}
(colored circles), in which the connecting line between the \acs{CMIP}5 and
\acs{CMIP}6 \ac{MMM} is not parallel to the \acs{CMIP}5 emergent relationships
for all emergent constraints.

>>>>

The massive spread in \ac{ECS} within the \acs{CMIP}6 ensemble discussed in the
previous \lcnamecref{ch:04:papers_ecs_tcr_assessment} begs the question about
the interpretation and credibility of these results. Since the \acs{CMIP}6
models provide crucial input for the upcoming international climate assessment
of the \ac{IPCC} (\acs{AR}6), this is not only important for the scientific
community, but also directly affects policymakers and society. Due to the many
new and improved processes included to the models, the \acs{CMIP}6 ensemble is
arguably more realistic than its predecessor generations. However, does this
automatically make results derived from \acs{CMIP}6 models more trustworthy
than those derived from earlier \ac{CMIP} models? In general, the answer to
this question is \enquote{not necessarily}. Even the \acs{CMIP}6 models are far
from fully simulating all aspects and processes of the full vastly complex
Earth system. Therefore, the addition of new processes to these models could
introduce new biases which have not been present in earlier \ac{CMIP} models.
For example, the apparent increase in climate sensitivity in the \acs{CMIP}6
models might be wrongly caused by missing processes that cancel out the effects
of the newly included processes. Thus, it is vital to take other sources of
information into account to assess the credibility of \ac{ESM} results.

@@

\begin{table}[!t]
  \centering
  \hbadness=3500
  \begin{tabular}{p{0.07\columnwidth} p{0.24\columnwidth} p{0.29\columnwidth}
      p{0.29\columnwidth}}
    \toprule
    Label & Reference & Description of \xaxis{} & Variables and corresponding
    observations \\
    \midrule
    BRI & \textcite{Brient2016} & Sensitivity of shortwave cloud albedo
    to changes in \acf{SST} [$\si{\pctK{}}$] & \makecell{\tabitem\emph{hur}
      (ERA-Interim) \\ \tabitem \emph{rsdt} (CERES-EBAF) \\ \tabitem
      \emph{rsut} (CERES-EBAF) \\ \tabitem \emph{rsutcs} (CERES-EBAF) \\
      \tabitem \emph{ts} (HadISST)} \\
    \midrule
    COX & \textcite{Cox2018} & $\Psi$ (temperature variability metric)
    [$\si{K}$] &
    \tabitem \emph{tas} (HadCRUT4) \\
    \midrule
    LIP & \textcite{Lipat2017} & Extent of Southern Hemisphere Hadley cell
    [$\si{\degree}$] & \tabitem \emph{va} (ERA-Interim) \\
    \midrule
    SHD & \textcite{Sherwood2014} & $D$-index (large-scale lower-tropospheric
    mixing) [$\si{1}$] & \tabitem \emph{wap} (ERA-Interim) \\
    \midrule
    SHL & \textcite{Sherwood2014} & \acs{LTMI} (\acl{LTMI}) [$\si{1}$] &
    \makecell{\tabitem \emph{hur} (ERA-Interim) \\ \tabitem \emph{ta}
      (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \midrule
    SHS & \textcite{Sherwood2014} & $S$-index (small-scale lower-tropospheric
    mixing) [$\si{1}$] & \makecell{\tabitem \emph{hur} (ERA-Interim) \\
      \tabitem \emph{ta} (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \midrule
    SU & \textcite{Su2014} & Error in vertical profile of relative humidity
    [$\si{1}$] & \tabitem \emph{hur} (AIRS, MLS-Aura) \\
    \midrule
    TIH & \textcite{Tian2015} & Tropical mid-tropospheric humidity asymmetry
    index from AIRS [$\si{\%}$] & \tabitem \emph{hus} (AIRS) \\
    \midrule
    TII & \textcite{Tian2015} & Southern \acs{ITCZ} index from GPCP
    [$\si{\mmday{}}$] & \tabitem \emph{pr} (GPCP) \\
    \midrule
    VOL* & \textcite{Volodin2008} & Difference between tropical and southern
    midlatitudinal cloud fraction [$\si{\%}$] & \tabitem \emph{clt} (ISCCP
    D-2) \\
    \midrule
    ZHA* & \textcite{Zhai2015} & Seasonal response of \acl{MBLC} fraction to
    changes in \acs{SST} [$\si{\pctK{}}$] &
    \makecell{\tabitem \emph{cl} (CloudSat/CALIPSO) \\ \tabitem \emph{tos}
      (AMSRE \acs{SST}) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \bottomrule
  \end{tabular}
  \caption[
    Overview of the 11 emergent constraints on the \acf{ECS} evaluated in
    \cref{ch:05:paper_ecs}.
  ]{
    Overview of the 11 emergent constraints on the \acf{ECS} evaluated in this
    \lcnamecref{ch:05:paper_ecs}. Detailed descriptions of the variables used
    to calculate the \xaxis{} values of the emergent constraints and the
    references for the corresponding observations (fourth column) are given in
    \cref{tab:app:a:overview_variables,tab:app:a:observations}, respectively.
    For emergent constraints marked with an asterisk (*), the observational
    value of the original publication is used to calculate the observational
    constraint. For all others, the observational value is computed with the
    \acs{ESMValTool}. \AdaptedFrom{Schlund2020a}.
  }
  \label{tab:05:overview_emergent_constraints}
\end{table}

This work is already published in \textcite{Schlund2020a}. For this paper, the
author of this thesis led the writing and the analysis and implemented the code
to reproduce this study with all figures and tables using the \ac{ESMValTool}.
In this \lcnamecref{ch:05:paper_ecs}, we first introduce the data (climate
model output and observations) and methods used in this study
(\cref{sec:05:data_and_methods}). After that, we present the 11 emergent
constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5 and \acs{CMIP}6
ensemble (\cref{sec:05:comparison_of_emergent_constraints}). Finally, we
provide a discussion (\cref{sec:05:discussion}) and a summary of the results
(\cref{sec:05:summary}).

@@

In this \lcnamecref{sec:05:comparison_of_emergent_constraints} we describe and
discuss the 11 emergent constraints on \ac{ECS} summarized in
\cref{tab:05:overview_emergent_constraints} using \acs{CMIP}5 and \acs{CMIP}6
data
(\cref{subsec:05:bri,subsec:05:cox,subsec:05:lip,subsec:05:shd,subsec:05:shs,%
  subsec:05:shl,subsec:05:su,subsec:05:tih,subsec:05:tii,subsec:05:vol,%
  subsec:05:zha}) and provide a best estimate for \ac{ECS} and statistical
significance of the 11 emergent constraints in
\cref{subsec:05:emergent_constraints_summary}. While most of these emergent
constraints have been derived using data from the \acs{CMIP}5 and/or
\acs{CMIP}3 ensembles, to our knowledge none of them has been evaluated on
the \acs{CMIP}6 ensemble so far. The results for the individual emergent
constraints described in the following are shown in
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}.
The left columns in these figures show the emergent relationships, including
the uncertainty of the linear regressions (blue and orange shaded areas; see
\cref{eq:02:spe}) and the uncertainty in the observations (gray shaded area;
see \cref{eq:02:pdf_obs}). The right columns show the \acp{PDF} of \ac{ECS} in
the original model ensemble (histogram) and the constrained distribution given
by the emergent constraints (blue and orange line; see
\cref{eq:02:pdf_y_given_x0}). \Cref{tab:05:overview_results} shows the
corresponding $66 \unit{\%}$ confidence intervals (\ie{} the
$\rangeunit{17}{83}{\%}$ intervals) of \ac{ECS} derived from the \acp{PDF}
given by \cref{eq:02:pdf_y_given_x0} and the $p$-values used to assess the
significance of the emergent relationships.

@@

As shown in the previous
\lcnamecrefs{sec:05:comparison_of_emergent_constraints}, most emergent
relationships show smaller coefficients of determination when evaluated on the
new \acs{CMIP}6 ensemble compared to the \acs{CMIP}5 ensemble. In this
\lcnamecref{sec:05:discussion}, we discuss possible reasons for, and
implications of, these differences. As reported by \textcite{Caldwell2014}, the
large amount of data provided by modern \acp{ESM} can generate spurious
correlations of variables between past climate and \ac{ECS} just by chance,
especially when only a few climate models are considered. This would cause the
performance of the emergent constraint to be reduced on out-of-sample data
(like the new \acs{CMIP}6 ensemble), since the emergent relationship appeared
just by chance and not because of a physically based mechanism.

@@

In this \lcnamecref{ch:05:paper_ecs}, 11 different emergent constraints on
\ac{ECS}, which are mostly directly or indirectly related to cloud feedbacks,
are assessed by applying them to results from \acp{ESM} contributing to
\acs{CMIP}5 and \acs{CMIP}6. Of particular interest are the results from
\acs{CMIP}6, since all analyzed emergent constraints have been published prior
to the availability of \acs{CMIP}6 data. In summary, the best estimate of
\ac{ECS} averaged over all emergent constraints increases by $12 \unit{\%}$
when moving from relationships trained on \acs{CMIP}5 to those trained on
\acs{CMIP}6. Some increase is predicted by almost every constraint we analyze
(with TII as the only exception) and can at least be partly explained by the
increased multi-model mean \ac{ECS} of \acs{CMIP}6, which has not been
accompanied by systematic changes in the constraint variables that could
explain this increase, leading to regression fits with higher intercept values
at observed constraint values. This is also illustrated by the \acs{CMIP}5 and
\acs{CMIP}6 \acp{MMM} in the left columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}
(colored circles), in which the connecting line between the \acs{CMIP}5 and
\acs{CMIP}6 \ac{MMM} is not parallel to the \acs{CMIP}5 emergent relationships
for all emergent constraints.
